---
title: "weisimu: Using Simulations to Assess the Performance of the Truncated Mean of a Sample Generated from a Weibull Distribution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trimmed-Mean-Weibull}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Pre-requisites

Install the package if needed and load it:

```{r loadpackage}
if (!require("devtools")) install.packages("devtools")
if (!require("weisimu")) devtools::install_github("Dorayaya/weisimu")
library(weisimu)
```

# Trimming the mean

We compare sample mean to trimmed mean from above. (GIVE FORMULAS).
For sample mean, know that it is exactly unbiased with variance (GIVE FORMULA).
So only need to simulate trimmed mean.

The base R `mean` function has a trim option, but trims tail above and below.
Does not allow to trim above only.
We add this functionality to the `mean` function as below:

```{r}
data = 1:100
mean(data,trim=0.05) # trims 2.5% above and 2.5% below
mean(data,trim=0.05,trim.upper=TRUE) # trims 5% above
```

# The Weibull Distribution

GIVE FORMULAS ETC. MAYBE SHOW DENSITY PLOT FROM SLIDES.


# Simulation Function

Here we simulate for some values of the parameters. The function `simtrim` spits out mean, bias, variance,
and MSE of sample mean and trimmed mean. As noted above, sample mean is always
unbiased. Variance in this case is 1, so MSE is 1. On the other hand,
trimmed mean trimed at 5\% introduces bias of about $-0.66$ but reduces
variance to about $0.4$. Thus overall MSE is about $0.85$, or 15\% smaller
than the unbiased gold standard.

```{r}
simtrim(n=20, shape=0.5, scale=1, p=0.05, S=10000)
```


# Simulating for ranges of values

We can run simulations for many values of a parameter using the `simtrim_by`
function. It takes the same arguments, but one is given as a vector.
It then runs S simulations for each value of the vector and returns the
results in a matrix. By default, the function will also plot the MSE of the
trimmed and untrimmed means, but you can turn this off with the option `plot=FALSE`.

```{r}
output = simtrim_by(n=seq(20,100,1),shape=0.5,scale=1,p=0.05,
                    S=1000,plot=FALSE)
output[,1:5]
```

The default plot:

```{r}
output = simtrim_by(n=seq(20,100,1),shape=0.5,scale=1,p=0.05,
                    S=1000)
```


We can also adjust the plot settings, including using syntax from the base `plot` function:

```{r}
output = simtrim_by(n=seq(20,100,1),shape=0.5,scale=1,p=0.05,
                    S=1000, leg.pos="bottomleft",
                    lty=c(1,2),lwd=c(2,3),col=c("purple","green"),
                    main="An Interesting Title")
```

## Adjusting shape param

As $\beta \to 0$ the mean and variance of weibull go to $\infty$.
We thus see that as $\beta \to 0$, the trimmed mean becomes more and more useful:

```{r}
output = simtrim_by(n=seq(20,100,1),shape=0.5,scale=1,p=0.05,
                    S=1000,
                    leg.pos="topright",col=c("red","blue"))
```

```{r}
output = simtrim_by(n=seq(20,100,1),shape=0.4,scale=1,p=0.05,
                    S=1000,
                    leg.pos="topright",col=c("red","blue"))
```


```{r}
output = simtrim_by(n=seq(20,100,1),shape=0.3,scale=1,p=0.05,
                    S=1000,
                    leg.pos="topright",col=c("red","blue"))
```

```{r}
output = simtrim_by(n=seq(20,500,10),shape=0.2,scale=1,p=0.05,
                    S=1000,
                    leg.pos="topright",col=c("red","blue"))
```



# Adjusting Trimming Proportion

We see that for a sample size of 300 and $\beta = 0.2$,
MSE favors only a little trimming--that is, only the largest observation.
Once the trimming proportion is around 0.05 or greater, the sample mean
becomes better.

```{r}
output = simtrim_by(n=300,shape=0.2,scale=1,p=seq(0.005,0.15,0.005),
                    S=2000,
                    leg.pos="bottomright",col=c("red","blue"))
```





